Enhancing Adversarial Robustness for Deep Metric Learning

### 2021-10-26 robdml.tex
tex

As pushing $\tilde{H}_\mathsf{S}$ towards the upperbound will easily lead to
model collapse, a valid $H_D\mathsf{D}$ should be choosed within the interval
$[H_\mathsf{S},2+\gamma]$ without rendering model collapse.
%
Thus, intuitively, $H_\mathsf{D}$ can be the hardness of another benign triplet
(with the same anchor) sampled with a strategy with a higher hardness
expectation, \ie, $E[H_\mathsf{D}] > E[H_\mathsf{S}]$.
%
Or at least the $\text{Var}[H_\mathsf{D}]$ of such sampling strategy has to be
large enough in order to create valid adversarial examples.
%
For instance, $H_\mathsf{D}$ can be the hardness of a Semihard~\cite{facenet}
triplet when the \emph{source} triplet is sampled with Random sampler.
%
Predictably, the model performance will be significantly influenced by the
triplet sampling strategies we chose for $H_\mathsf{D}$ in this case.
%
For convenience of further discussion, we denote the hardness of a Random
triplet, Semihard triplet, and Softhard triplet as $\mathcal{R}$,
$\mathcal{M}$, $\mathcal{S}$, respectively.

If we have a strong prior knowledge on what the \emph{destination} hardness
should be, we can even use a pseudo-hardness function $g(\cdot)$, which is a
customized scalar function.

@@

\cref{fig:ga}

@@

When benign triplets are used as reference for destination hardness, the
adversary (\ie, hardness of adversarial triplets) could gradually become weaker
than it should be in the late phase of training.
%
During the training process of a DML model, the expectation of the triplet loss
$E[L_\text{T}(\cdots)]$ on benign examples is gradually decreasing
(with the lower bound being zero).
%
The corresponding expectation of hardness $E[H(\cdots)]$ will
decrease as well (with the lower bound being $-\gamma$).
%
As a result, $E[H_\mathsf{D}-H_\mathsf{S}]$ will gradually tend to zero,
indicating a gradually decreasing strength of attack, and inefficiency in
learning from a strong adversary.

% hardness boosting

>>>>

As pushing $\tilde{H}_\mathsf{S}$ towards the upperbound will easily lead to
model collapse, a valid $H_\mathsf{D}$ should be choosed within the interval
$[H_\mathsf{S},2+\gamma]$ without rendering model collapse.
%
Thus, intuitively, $H_\mathsf{D}$ can be the hardness of another benign triplet
(with the same anchor) sampled with a strategy with a higher hardness
expectation, \ie, $E[H_\mathsf{D}] > E[H_\mathsf{S}]$.
%
Or at least the $\text{Var}[H_\mathsf{D}]$ of such sampling strategy has to be
large enough in order to create valid adversarial examples.
%
For instance, $H_\mathsf{D}$ can be the hardness of a Semihard~\cite{facenet}
triplet when the \emph{source} triplet is sampled with Random sampler.
%
Predictably, the model performance will be significantly influenced by the
triplet sampling strategies we chose for $H_\mathsf{D}$ in this case.
%
For convenience of further discussion, we denote the hardness of a Random
triplet, Semihard triplet, and Softhard triplet as $\mathcal{R}$,
$\mathcal{M}$, $\mathcal{S}$, respectively.

If we have a strong prior knowledge on what the \emph{destination} hardness
should be, we can even use a pseudo-hardness function $g(\cdot)$, \ie, a
customized scalar function.

@@

@@

Even if $H_\mathsf{D}$ is calucated from another triplet harder than the
\emph{source} triplet, the adversary may become weak in the late phase of
training.
%
The training process of the model aims to reduce the expectation of triplet loss
$E[L_\text{T}]$ to zero over the distribution of sample triplets,
and thus the $E[H]$ of any given triplet will eventually tend to $-\gamma$,
reducing the hardness of adversarial triplets from HM.

\cref{fig:ga}
% hardness boosting
